Governing through patronage: the bargain for education in decentralized Brazil

### 2020-11-13 paper/paper.tex
fix width of graphics

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\renewcommand{\theproposition}{\arabic{proposition}}
\newcommand{\real}{\mathbb{R}_+^n}
\newcommand{\bfv}{\mathbf{v}}
\newcommand{\blue}{\textcolor{blue}}
% \renewcommand{\includegraphics}{\includegraphics[width=0.6\linewidth]}

@@

Brazil's local level of government is composed of over 5 thousand municipalities, each with their own mayorship (\emph{prefeitura}) and city council (\emph{c\^{a}mara dos vereadores}). Mayors and city councillors (\textit{vereadores}) are democratically elected every four years, open to reelection for both offices.\footnote{The last municipal elections were in 2016. These take place every 4 years.} The executive is responsible for the management of public services such as education, with exclusive rights over public administration in general, and personnel appointment in particular \citep{akhtari_political_2015,brollo_victor_2017}. As noted by \citet{souza_reforma_2004}, decisions over whom to appoint into the educational sector are primarily in the hands of the mayor and her secretary of education. The city council, on the other hand, oversees legislation and approves the budget for the fiscal year, providing a local checks and balances on the executive.

At the federal level, this instituitonal arrangement gives rise to a phenomenon dscribed as presidential coalitionism \citet{raile_executive_2011, power_optimism_2010}. Executive leaders garner legislative support from the National Congress by exchanging key positions in the federal bureaucracy, appointing members of their legislative coalition into cabinet positions. \citep{raile_executive_2011}. In a setting characterized by weak party cohesion and programmatic commitments \citet{ames_electoral_1995, lucas_ideological_2010}, bureaucratic positions for members of the coalition provide material incentives for legislators to support the executive agenda \citet{batista_o_2013, neto_presidential_2006, figueiredo_executivo_1999}.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/descriptive_enrollment.pdf}
    \caption{\textbf{Student enrollment by administration.}}
    \label{fig:enroll}
\end{figure}

@@

\begin{table}[h]
    \centering
    \footnotesize
    \input{results/turnover_result}
    \caption{{\bf Executive coalitions and staff turnover.} An increase in the share of legislative seats held by the mayoral coalition decrease the amount of turnover for teachers and school principals, including hires or dismissals. Models 1 and 2 present results on the turnover index at the school level. Models 2 and 3 are aggregate hiring rates at the municipal level. Models 5 and 6 are logistic regressions at the individual, bureaucrat level. where the outcome of interest is the proportion of staff either hired or dismissed at a given year. Models 1, 3, and 5 include year and state fixed effects.}
    \label{tbl:turnover}
  \end{table}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/descriptive_staff_breakdown}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\linewidth]{plots/descriptive_staff_turnover.pdf}
    \caption{Proportion of public sector jobs by department. Note that I only keep the top 5 categorie}
    \label{fig:staff_turnover}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{plots/hist_coalition.pdf}
    \caption{Distribution of share of legislative seats allied with the mayor during the electoral campaign for the elecoral period of 2002-2016.}
    \label{fig:hist_coalition}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_coalition_coef_mun.pdf}
    \caption{{\bf Political alignment and staff turnover.} A visual representation of results presented in table \ref{tbl:turnover}. All models include year and state fixed effects.}
    \label{fig:coalition_turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_interactions_logit}
    \caption{\textbf{Differential reductions in patronage in teachers and school principals}. Predicted values for bureaucratic turnover plotted against the proportion of educational staff either hired or fired for any given year against the share of seats controlled by the executive.}
    \label{fig:turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/plot_coalition_coef_mun_second.pdf}
    \caption{\textbf{Patronage continues into the second term.} There is strong evidence that executive-legislative bargains continue into the mayor's second term. All models include a full set of controls, as well as year and state fixed effects.}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\linewidth]{plots/model_turnover_learning.pdf}
  \caption{{\bf Bureaucratic turnover and student learning} Teacher and school principal turnover have a negative effect on student learning, across different sets of exams (\emph{SAEB} and \emph{SPAECE}). All models include a full set of controls, as well as year and state fixed effects.}
  \label{fig:hlm_mods}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\linewidth]{plots/reelection_marginal_effect_plots.pdf}
    \caption{\textbf{Differential returns to patronage}. Mayors who engage in patronage do not directly benefit from it. Instead, city councilors in the mayoral ally base seem to be benefiting from it at the expense of non-allied city councilors.}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\linewidth]{plots/visual_rdd_legislative_seat_flip.pdf}
\end{figure}

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\renewcommand{\theproposition}{\arabic{proposition}}
\newcommand{\real}{\mathbb{R}_+^n}
\newcommand{\bfv}{\mathbf{v}}
\newcommand{\blue}{\textcolor{blue}}
% \renewcommand{\includegraphics}{\includegraphics}

@@

Brazil's local level of government is composed of over 5 thousand municipalities, each with their own city hall (\emph{prefeitura}) and city council (\emph{c\^{a}mara dos vereadores}). Mayors and city councilors (\textit{vereadores}) are democratically elected every four years, open to reelection for both offices.\footnote{The last municipal elections were in 2016. These take place every 4 years.} The executive is responsible for the management of public services such as education, with exclusive rights over public administration in general, and personnel appointment in particular \citep{akhtari_political_2015,brollo_victor_2017}. As noted by \citet{souza_reforma_2004}, decisions over whom to appoint into the educational sector are primarily in the hands of the mayor and her secretary of education. The city council, on the other hand, oversees legislation and approves the budget for the fiscal year, providing a local checks and balances on the executive.

At the federal level, this institutional arrangement gives rise to a phenomenon described as presidential coalitionism \citet{raile_executive_2011, power_optimism_2010}. Executive leaders garner legislative support from the National Congress by exchanging key positions in the federal bureaucracy, appointing members of their legislative coalition into cabinet positions. \citep{raile_executive_2011}. In a setting characterized by weak party cohesion and programmatic commitments \citet{ames_electoral_1995, lucas_ideological_2010}, bureaucratic positions for members of the coalition provide material incentives for legislators to support the executive agenda \citet{batista_o_2013, neto_presidential_2006, figueiredo_executivo_1999}.

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/descriptive_enrollment.pdf}
    \caption{\textbf{Student enrollment by administration.}}
    \label{fig:enroll}
\end{figure}

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/descriptive_staff_breakdown}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/descriptive_staff_turnover.pdf}
    \caption{Proportion of public sector jobs by department. Note that I only keep the top 5 categorie}
    \label{fig:staff_turnover}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{plots/hist_coalition.pdf}
    \caption{Distribution of share of legislative seats allied with the mayor during the electoral campaign for the elecoral period of 2002-2016.}
    \label{fig:hist_coalition}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/plot_coalition_coef_mun.pdf}
    \caption{{\bf Political alignment and staff turnover.} A visual representation of results presented in table \ref{tbl:turnover}. All models include year and state fixed effects.}
    \label{fig:coalition_turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/plot_interactions_logit}
    \caption{\textbf{Differential reductions in patronage in teachers and school principals}. Predicted values for bureaucratic turnover plotted against the proportion of educational staff either hired or fired for any given year against the share of seats controlled by the executive.}
    \label{fig:turnover}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/plot_coalition_coef_mun_second.pdf}
    \caption{\textbf{Patronage continues into the second term.} There is strong evidence that executive-legislative bargains continue into the mayor's second term. All models include a full set of controls, as well as year and state fixed effects.}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics{plots/model_turnover_learning.pdf}
  \caption{{\bf Bureaucratic turnover and student learning} Teacher and school principal turnover have a negative effect on student learning, across different sets of exams (\emph{SAEB} and \emph{SPAECE}). All models include a full set of controls, as well as year and state fixed effects.}
  \label{fig:hlm_mods}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/reelection_marginal_effect_plots.pdf}
    \caption{\textbf{Differential returns to patronage}. Mayors who engage in patronage do not directly benefit from it. Instead, city councilors in the mayoral ally base seem to be benefiting from it at the expense of non-allied city councilors.}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics{plots/visual_rdd_legislative_seat_flip.pdf}
\end{figure}
